Self Referencing Sequences

### 2014-11-05 draft.tex
Restructuring Equivalence section

In this section, we consider long term equivalence of sequences. We introduce a definition for this
\begin{definition}
We say that two sequence $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n, k \geq 0$ such that 
\begin{align*}
a_{n+i} = b_{k+i}
\end{align*}
for all positive integers $i$. 
\end{definition}

 Consider, for example, sequences generated by $C = (123)$. Theorem~\ref{independentsequences} shows that the sequence starting with 2
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
and the sequence starting with 3
\begin{align*}
3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
are independent in the sense that neither sequence is exactly contained in the other. However, if we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we get 
\begin{align*}
2,2,|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots\\
3,|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
in both cases.

We now introduce a notion which will help us categorize sequences as equivalent. 
\begin{definition} Consider a sequence $\{a_i\}$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
\end{definition}
The read distance can be any nonnegative integer. For example if we consider the sequence $\{a_i\}$ generated by $C = (1 2 3)$ beginning with a 1  
\begin{align*}
1,2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
the read distance for $1$ is always $0$ since in a self-referencing sequence, the first number must refer to the first block. The read distance for $2$ is $0$ as the second block of numbers starts at the second number. We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0, 0 ,1,2,4,6,6,6,6,7,8,9,\ldots$. The read distance, can be thought of as the horizontal shift of the arrows in Figure~\ref{fig:onetwothree}.

\begin{lemma} 
\label{readdistformula}
Let $\{a_i\}$ be any self referencing sequence generated by some set of positive integers. Then the read distance for $i>1$, is 
\begin{equation*}
\sum_{k=1}^{i-1}( a_k - 1)
\end{equation*} 
and the read distance for $1$ is always 0.
\end{lemma}
\begin{proof}
We will prove this using induction. We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number. When $i=2$ the read distance is $a_1-1$. The first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at position $a_1$. The number of elements between position $a_1$ and $2$ is exactly $a_1 -1$.

Now if we assume that our result holds for $i=l$ we would like to show that it holds for $i=l+1$. We know that the distance between the start of the $l$th block and position $l$ is 
\begin{equation*}
\sum_{k=1}^{l-1}( a_k - 1).
\end{equation*} 
This $l$th block has length $a_l$ and so the $l+1$th block will begin $a_l$ positions to the right of the beginning of the $l$th block. Also note that $a_{l+1}$ is one position to the right of $a_l$. Therefore the beginning of the $(l+1)$th block is 
\begin{equation*}
a_l-1 + \sum_{k=1}^{l-1} (a_k-1) = \sum_{k=1}^{l} (a_k-1)
\end{equation*} 
positions to the right of $a_{l+1}$. But this is exactly the reading distance for $a_{l+1}$ and so we have our result.
\end{proof}
Note that because $a_i$ is generated by positive integers we always have $a_k-1\geq 0$ and so the read distance is monotonically increasing. Although this shows us that read distance never decreases it does not tell us how quickly it increases, our next result gives a simple bound on the number of consecutive times the read distance can stay constant.

\begin{lemma} Consider a self referencing sequence $\{a_i\}$ generated by $(c_1 c_2 \ldots c_n)$ with $m = \max(c_1, c_2, \ldots, c_n)$. Suppose the read distance for $i$ is $d$, then the read distance for $i+m$ is at least $d+1$. 
\end{lemma}
\begin{proof}
The read distances can stay constant from $i$ to $i+1$ only if there is a $a_{i+1} = 1$. There can be at most $m$ 1's in a contiguous block, so the read distance of $i+m$ is bigger than the read distance of $i$.  
\end{proof}

\begin{theorem} If two sequences are equivalent then the first element of the shared sequence must have the same read distance in both sequences.
\end{theorem}
\begin{proof}
Assume otherwise. Suppose the two sequences are $\{a_j\}$ and $\{b_j\}$ and we have 
\begin{equation*}
a_{i+l} = b_{i+k} = c_{i}
\end{equation*}
for all $i>1$, such that if $m$ or $l$ were made smaller this relation would not hold. Suppose the read distance for $b_{k+1}$ is $x$ and the read distance for $a_{1+l}$ is $y$. Without loss of generality let us assume that $x>y$ (note that $x \neq y$ as per our assumption).

We also have a sufficient condition for equivalence 
\begin{proposition} 
\label{equivsuff}
If two sequences $\{a_i\}, \{b_i\}$ are such that for some integers $m,l$ the read distance for $a_m$ and $b_l$ are both $d$, some integer and 
\begin{equation*}
a_{m+k} = b_{l+k}
\end{equation*}
where $0 \leq k \leq d-1$ then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
\end{proposition}
\begin{proof}
We will prove this by  induction. In particular we will show for all $k \geq 0$ that $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$. By our assumptions we already have this for $0 \leq k \leq d-1$ and so we begin with $k=d$.

@@

This result allows us to make a general statement for equivalence for sequences that begin with a $1$.
\begin{corollary} If $C = (1 c_1 c_2 \cdots c_n)$, then $s(C, 1)$ and $s(C, c_1)$ are equivalent.
\end{corollary}
\begin{proof}
Let $\{a_i\} = s(C, 1)$ and $\{b_i\} = s(C, c_1)$. The read distance of sequence $\{a_i\}$ for $i=2$ is 0, and the read distance of sequence $\{b_i\}$ for $i = 1$ is 0. In addition, $a_2 = b_1 = c_1$, so by Proposition ~\ref{equivsuff} the two sequences are equivalent.
\end{proof}

>>>>

In this section, we consider long term equivalence of sequences. We introduce a formal definition of this and establish results that allow us to determine when two sequences are equivalent.

We would first like to make a formal definition of equivalence. 
\begin{definition}
We say that two sequence $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n, k \geq 0$ such that 
\begin{align*}
a_{n+i} = b_{k+i}
\end{align*}
for all positive integers $i$. 
\end{definition}

As an example let us consider sequences generated by $C = (123)$. The sequence which starts with a 2 is 
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
\end{align*}
and the sequence starting with 3 is 
\begin{align*}
3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ .
\end{align*}
If we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we see 
\begin{align*}
2,2,&|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots\\
3,&|3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots \ ,
\end{align*}
that the resultant sequences appear to be identical and thus equivalent. At an intuitive level it seems obvious that these two truncated sequences are identical as they both begin with the same two numbers and the numbers which were removed do not refer to blocks to the right of these numbers.

In order to formalize and generalize this idea we introduce the notion of a read distance.  
\begin{definition} Consider a sequence $\{a_i\}$. We define the read distance for an index $i$ as the number of terms in between it $a_i$ and the beginning of the $i$th block of numbers.
\end{definition}
The read distance can be any nonnegative integer. For example if we consider the sequence $\{a_i\}$ generated by $C = (1 2 3)$ beginning with a 1  
\begin{align*}
1,2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
the read distance for $1$ is always $0$ since in a self-referencing sequence, the first number must refer to the first block. The read distance for $2$ is $0$ as the second block of numbers starts at the second number. We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0, 0 ,1,2,4,6,6,6,6,7,8,9,\ldots$.

This monotonicity is true in general, and we can in fact make a stronger statement on the value of the read distance at a given index.
\begin{lemma} 
\label{readdistformula}
Let $\{a_i\}$ be any self referencing sequence generated by some set of positive integers. Then the read distance for $i>1$, is 
\begin{equation*}
\sum_{k=1}^{i-1}( a_k - 1)
\end{equation*} 
and the read distance for $1$ is always 0.
\end{lemma}
\begin{proof}
We will prove this using induction. We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number. When $i=2$ the read distance is $a_1-1$. The first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at position $a_1$. The number of elements between position $a_1$ and $2$ is exactly $a_1 -1$.

Now if we assume that our result holds for $i=l$ we would like to show that it holds for $i=l+1$. We know that the distance between the start of the $l$th block and position $l$ is 
\begin{equation*}
\sum_{k=1}^{l-1}( a_k - 1).
\end{equation*} 
This $l$th block has length $a_l$ and so the $l+1$th block will begin $a_l$ positions to the right of the beginning of the $l$th block. Also note that $a_{l+1}$ is one position to the right of $a_l$. Therefore the beginning of the $(l+1)$th block is 
\begin{equation*}
a_l-1 + \sum_{k=1}^{l-1} (a_k-1) = \sum_{k=1}^{l} (a_k-1)
\end{equation*} 
positions to the right of $a_{l+1}$. But this is exactly the reading distance for $a_{l+1}$ and so we have our result.
\end{proof}
Note that because $a_i$ is generated by positive integers we always have $a_k-1\geq 0$ and so the read distance is monotonically increasing.

%Although this shows us that read distance never decreases it does not tell us how quickly it increases, our next result gives a simple bound on the number of consecutive times the read distance can stay constant. 
%\begin{corollary} Consider a self referencing sequence $\{a_i\}$ generated by $(c_1 c_2 \ldots c_n)$ with $m = \max(c_1, c_2, \ldots, c_n)$. Suppose the read distance for $i$ is $d$, then the read distance for $i+m$ is at least $d+1$. 
%\end{corollary}
%\begin{proof}
%The read distances can stay constant from $i$ to $i+1$ only if there is a $a_{i+1} = 1$. There can be at most $m$ 1's in a contiguous block, so the read distance of $i+m$ is bigger than the read distance of $i$.  
%\end{proof}

Now that we have a nice expression for the value of the read distance at a particular index we can show our first major result which allows us to check only finitely many indices in order to establish that two sequences are equivalent. 
\begin{theorem} 
\label{equivsuff}
If two sequences $\{a_i\}, \{b_i\}$ generated over the same generating cycle are such that for some integers $m,l$ the read distance for $a_m$ and $b_l$ are both $d$, some integer and 
\begin{equation*}
a_{m+k} = b_{l+k}
\end{equation*}
where $0 \leq k \leq d-1$ then $\{a_i\}$ and $\{b_i\}$ are equivalent. 
\end{theorem}
\begin{proof}
We will prove this by  induction. In particular we will show for all $k \geq 0$ that $a_{m+k} = b_{l+k}$ and the read distance for $a_{m+k}$ equals the read distance for $b_{l+k}$. By our assumptions we already have this for $0 \leq k \leq d-1$ and so we begin with $k=d$.

@@

This result also allows us to make a general statement for equivalence for sequences that begin with a $1$.
\begin{corollary} If $C = (1 c_1 c_2 \cdots c_n)$, then $s(C, 1)$ and $s(C, c_1)$ are equivalent.
\end{corollary}
\begin{proof}
Let $\{a_i\} = s(C, 1)$ and $\{b_i\} = s(C, c_1)$. The read distance of sequence $\{a_i\}$ for $i=2$ is 0, and the read distance of sequence $\{b_i\}$ for $i = 1$ is 0. In addition, $a_2 = b_1 = c_1$, so by Proposition ~\ref{equivsuff} the two sequences are equivalent.
\end{proof}

We also can demonstrate a sufficient condition for equivalence. 
\begin{theorem} If two sequences are equivalent then the first element of the shared sequence must have the same read distance in both sequences.
\end{theorem}
\begin{proof}
Assume otherwise. Suppose the two sequences are $\{a_j\}$ and $\{b_j\}$ and we have 
\begin{equation*}
a_{i+l} = b_{i+k} = c_{i}
\end{equation*}
for all $i>1$, such that if $m$ or $l$ were made smaller this relation would not hold. Suppose the read distance for $b_{k+1}$ is $x$ and the read distance for $a_{1+l}$ is $y$. Without loss of generality let us assume that $x>y$ (note that $x \neq y$ as per our assumption).
